---
title: "Analysis for explanatory language trial"
output: html_document
---

### Load necessary libraries for analysis

```{r}
library(readr)
library(dplyr)
library(devtools)
```

### Read in the de-identified .csv files for the original (results1) and replicated (results2) experiments

```{r}
results1 <- read_csv("dataanalysis-001 Quiz 1 responses_092916_deidentified.csv")
results2 <- read_csv("dataanalysis-002 Quiz 1 responses_092916_deidentified.csv")
```

### Clean data

The function *checkAndCleanData* does the following:

* Selects and renames the appropriate quiz questions for our experiment
* Subsets to only the first submission (lowest *submission_id*) for each student (*session_user_id*)
* Creates a variable denoting treatment group (*prompt*: clang, noclang)
* Removes students with missing data (those who didn't answer our question)
* Creates a variable denoting possible answer choices (*answerOptions*: IDCP, IDCM, IDPM, ICPM)
* Creates a variable denoting student's answer choice (*answerChoice*: inferential, descriptive, causal, predictive, mechanistic)
* Returns resulting clean data frame (*cleanResults*)

```{r}
.miss <- function(vec) {
    vec=="0" | vec=="NULL"
}

checkAndCleanData <- function(results) {
    # Select the columns pertaining to the experimental quiz question
    whCols <- grep("We take a random sample of individuals", results[1,])

    # Rename columns
    newColnames <- rep("noclang", length(whCols))
    newColnames[grep("benzene", results[1,whCols])] <- "clang"
    answerChoices <- gsub("This is an example of a(n)* ", "", results[2,whCols])
    answerChoices <- gsub(" data analysis(.)*", "", answerChoices)
    newColnames <- paste0(newColnames, "_", answerChoices)
    colnames(results)[whCols] <- newColnames

    rowStartResults <- which.max(results[,1] != "session_user_id")
    
    #Keep session_user_id, submission_id, and  questions
    results <- results[rowStartResults:nrow(results), c(1,4,whCols)]

    colsNoclang <- grep("^noclang_", colnames(results))
    colsClang <- grep("^clang_", colnames(results))
    checkMat <- cbind(rowSums(!.miss(results[,colsNoclang])),
        rowSums(!.miss(results[,colsClang])))

    # There should be no 2's (this would indicate a student answering both versions in the same quiz attempt)
    # 0's indicate that the student did not check a response
    print("Checking if students answered both versions in the same quiz attempt (no 2's)")
    print(table(rowSums(checkMat)))

    # Get the first submission for each session_user_id
    results <- results %>% subset(submission_id=="1")

    print("Should be 0 or 1...")
    print(table(rowSums(results[,colsNoclang]=="1")))
    print("Should be 0 or 1...")
    print(table(rowSums(results[,colsClang]=="1")))
    results$sawNoclang <- rowSums(results[,colsNoclang]=="1")==1
    results$sawClang <- rowSums(results[,colsClang]=="1")==1
    print("There should be no counts in TRUE-TRUE")
    print(table(sawClang = results$sawClang, sawNoclang = results$sawNoclang))

    # Only keep rows where a student answered question
    results <- results[results$sawClang | results$sawNoclang,]
    results <- results[!is.na(results$sawClang) & !is.na(results$sawNoclang),]
    
    # Create one variable for which prompt was seen
    results$prompt <- factor(ifelse(results$sawClang, "clang", "noclang"))

    # Get answer choices picked by the students in one variable
    answerChoiceCol <- sapply(1:nrow(results), function(r) {
        which(results[r,c(colsNoclang, colsClang)]=="1")
    })
    results$answerChoice <- factor(answerChoices[answerChoiceCol])

    #Get answer options that the student saw
    #I=inferential (1), D=descriptive (2), C=causal (3), P=predictive (4), M=mechanistic (5)
    #Possibilities are IDCP, IDCM, IDPM, ICPM
    answerOptions <- rep(NA, nrow(results))
    temp3 <- sapply(1:nrow(results), function(r) {
        temp4 = which(results[r,c(colsNoclang, colsClang)]!="NULL")
        #paste(temp4,collapse="")
        substr(paste(temp4,collapse=""),1,4)
    })
    answerOptions[temp3 == "1234"] <- "IDCP"
    answerOptions[temp3 == "1235"] <- "IDCM"
    answerOptions[temp3 == "1245"] <- "IDPM"
    answerOptions[temp3 == "1345"] <- "ICPM"
    
    table(answerOptions)
    
    results$answerOptions <- as.factor(answerOptions)
    
    cleanResults = results
    cleanResults
}
```

The function *getSubsetTable* does the following:

* Tables the clean results (*cleanResults*) by the answer choice (*answerChoice*) and treatment group (*prompt*) for the designated subset of answer choices (*answerOptions*: IDCP, IDCM, IDPM, ICPM)

```{r}
getSubsetTable <- function(cleanResults, subset) {   
    cleanResultsSubset <- cleanResults %>% filter(answerOptions==subset)
    table(cleanResultsSubset$answerChoice, cleanResultsSubset$prompt)
}
```

### Get clean data

```{r}
cleanResults1 <- checkAndCleanData(results1)  ## January 2013 course
cleanResults2 <- checkAndCleanData(results2)  ## October 2013 course
```


### Results for answer choices: inferential, descriptive, causal, predictive

Summarized student results (counts and percentages) for each answer choice (January 2013 course):

```{r}
## January 2013 course

tab1_IDCP <- getSubsetTable(cleanResults1, "IDCP")
addmargins(tab1_IDCP)
proptab1_IDCP <- prop.table(tab1_IDCP, margin=2)
round(proptab1_IDCP*100,1)
```

Total percentage of students who correctly answered "inferential" (January 2013 course):

```{r}
## January 2013 course
inferential1_IDCP = tab1_IDCP["inferential",]
totals1_IDCP = addmargins(tab1_IDCP)["Sum",][1:2]
sum(inferential1_IDCP)/sum(totals1_IDCP)
```

Summarized student results (counts and percentages) for each answer choice (October 2013 course):

```{r}
## October 2013 course
tab2_IDCP <- getSubsetTable(cleanResults2, "IDCP")
addmargins(tab2_IDCP)
proptab2_IDCP <- prop.table(tab2_IDCP, margin=2)
round(proptab2_IDCP*100,1)
```

Total percentage of students who correctly answered "inferential" (October 2013 course):

```{r}
## October 2013 course
inferential2_IDCP = tab2_IDCP["inferential",]
totals2_IDCP = addmargins(tab2_IDCP)["Sum",][1:2]
sum(inferential2_IDCP)/sum(totals2_IDCP)
```

Results of a two-proportions test, comparing the proportion of students who claimed analysis was "causal" between the explanatory language (*clang*) and non-explanatory language (*noclang*) treatment groups (January 2013 course):

```{r}
## January 2013 course
causal1_IDCP = tab1_IDCP["causal",]
totals1_IDCP = addmargins(tab1_IDCP)["Sum",][1:2]
prop.test(causal1_IDCP, totals1_IDCP)
```

Results of a two-proportions test, comparing the proportion of students who claimed analysis was "causal" between the explanatory language (*clang*) and non-explanatory language (*noclang*) treatment groups (October 2013 course):

```{r}
## October 2013 course
causal2_IDCP = tab2_IDCP["causal",]
totals2_IDCP = addmargins(tab2_IDCP)["Sum",][1:2]
prop.test(causal2_IDCP, totals2_IDCP)
```

### Results for answer choices: inferential, descriptive, causal, mechanistic

Summarized student results (counts and percentages) for each answer choice (January 2013 course):

```{r}
## January 2013 course

tab1_IDCM <- getSubsetTable(cleanResults1, "IDCM")
addmargins(tab1_IDCM)
proptab1_IDCM <- prop.table(tab1_IDCM, margin=2)
round(proptab1_IDCM*100,1)
```

Total percentage of students who correctly answered "inferential" (January 2013 course):

```{r}
## January 2013 course
inferential1_IDCM = tab1_IDCM["inferential",]
totals1_IDCM = addmargins(tab1_IDCM)["Sum",][1:2]
sum(inferential1_IDCM)/sum(totals1_IDCM)
```

Summarized student results (counts and percentages) for each answer choice (October 2013 course):

```{r}
## October 2013 course
tab2_IDCM <- getSubsetTable(cleanResults2, "IDCM")
addmargins(tab2_IDCM)
proptab2_IDCM <- prop.table(tab2_IDCM, margin=2)
round(proptab2_IDCM*100,1)
```

Total percentage of students who correctly answered "inferential" (October 2013 course):

```{r}
## October 2013 course
inferential2_IDCM = tab2_IDCM["inferential",]
totals2_IDCM = addmargins(tab2_IDCM)["Sum",][1:2]
sum(inferential2_IDCM)/sum(totals2_IDCM)
```

Results of a two-proportions test, comparing the proportion of students who claimed analysis was "causal" between the explanatory language (*clang*) and non-explanatory language (*noclang*) treatment groups (January 2013 course):

```{r}
## January 2013 course
causal1_IDCM = tab1_IDCM["causal",]
totals1_IDCM = addmargins(tab1_IDCM)["Sum",][1:2]
prop.test(causal1_IDCM, totals1_IDCM)
```

Results of a two-proportions test, comparing the proportion of students who claimed analysis was "causal" between the explanatory language (*clang*) and non-explanatory language (*noclang*) treatment groups (October 2013 course):

```{r}
## October 2013 course
causal2_IDCM = tab2_IDCM["causal",]
totals2_IDCM = addmargins(tab2_IDCM)["Sum",][1:2]
prop.test(causal2_IDCM, totals2_IDCM)
```


### Results for answer choices: inferential, descriptive, predictive, mechanistic

Summarized student results (counts and percentages) for each answer choice (January 2013 course):

```{r}
## January 2013 course

tab1_IDPM <- getSubsetTable(cleanResults1, "IDPM")
addmargins(tab1_IDPM)
proptab1_IDPM <- prop.table(tab1_IDPM, margin=2)
round(proptab1_IDPM*100,1)
```

Total percentage of students who correctly answered "inferential" (January 2013 course):

```{r}
## January 2013 course
inferential1_IDPM = tab1_IDPM["inferential",]
totals1_IDPM = addmargins(tab1_IDPM)["Sum",][1:2]
sum(inferential1_IDPM)/sum(totals1_IDPM)
```

Summarized student results (counts and percentages) for each answer choice (October 2013 course):

```{r}
## October 2013 course
tab2_IDPM <- getSubsetTable(cleanResults2, "IDPM")
addmargins(tab2_IDPM)
proptab2_IDPM <- prop.table(tab2_IDPM, margin=2)
round(proptab2_IDPM*100,1)
```

Total percentage of students who correctly answered "inferential" (October 2013 course):

```{r}
## October 2013 course
inferential2_IDPM = tab2_IDPM["inferential",]
totals2_IDPM = addmargins(tab2_IDPM)["Sum",][1:2]
sum(inferential2_IDPM)/sum(totals2_IDPM)
```

Results of a two-proportions test, comparing the proportion of students who claimed analysis was "inferential" between the explanatory language (*clang*) and non-explanatory language (*noclang*) treatment groups (January 2013 course):

```{r}
## January 2013 course
inferential1_IDPM = tab1_IDPM["inferential",]
totals1_IDPM = addmargins(tab1_IDPM)["Sum",][1:2]
prop.test(inferential1_IDPM, totals1_IDPM)
```

Results of a two-proportions test, comparing the proportion of students who claimed analysis was "inferential" between the explanatory language (*clang*) and non-explanatory language (*noclang*) treatment groups (October 2013 course):

```{r}
## October 2013 course
inferential2_IDPM = tab2_IDPM["inferential",]
totals2_IDPM = addmargins(tab2_IDPM)["Sum",][1:2]
prop.test(inferential2_IDPM, totals2_IDPM)
```

### Results for answer choices: inferential, causal, predictive, mechanistic

Summarized student results (counts and percentages) for each answer choice (January 2013 course):

```{r}
## January 2013 course

tab1_ICPM <- getSubsetTable(cleanResults1, "ICPM")
addmargins(tab1_ICPM)
proptab1_ICPM <- prop.table(tab1_ICPM, margin=2)
round(proptab1_ICPM*100,1)
```

Total percentage of students who correctly answered "inferential" (January 2013 course):

```{r}
## January 2013 course
inferential1_ICPM = tab1_ICPM["inferential",]
totals1_ICPM = addmargins(tab1_ICPM)["Sum",][1:2]
sum(inferential1_ICPM)/sum(totals1_ICPM)
```

Summarized student results (counts and percentages) for each answer choice (October 2013 course):

```{r}
## October 2013 course
tab2_ICPM <- getSubsetTable(cleanResults2, "ICPM")
addmargins(tab2_ICPM)
proptab2_ICPM <- prop.table(tab2_ICPM, margin=2)
round(proptab2_ICPM*100,1)
```

Total percentage of students who correctly answered "inferential" (October 2013 course):

```{r}
## October 2013 course
inferential2_ICPM = tab2_ICPM["inferential",]
totals2_ICPM = addmargins(tab2_ICPM)["Sum",][1:2]
sum(inferential2_ICPM)/sum(totals2_ICPM)
```

Results of a two-proportions test, comparing the proportion of students who claimed analysis was "causal" between the explanatory language (*clang*) and non-explanatory language (*noclang*) treatment groups (January 2013 course):

```{r}
## January 2013 course
causal1_ICPM = tab1_ICPM["causal",]
totals1_ICPM = addmargins(tab1_ICPM)["Sum",][1:2]
prop.test(causal1_ICPM, totals1_ICPM)
```

Results of a two-proportions test, comparing the proportion of students who claimed analysis was "mechanistic" between the explanatory language (*clang*) and non-explanatory language (*noclang*) treatment groups (January 2013 course):

```{r}
## January 2013 course
mech1_ICPM = tab1_ICPM["mechanistic",]
prop.test(mech1_ICPM, totals1_ICPM)
```

Results of a two-proportions test, comparing the proportion of students who claimed analysis was "causal" between the explanatory language (*clang*) and non-explanatory language (*noclang*) treatment groups (October 2013 course):

```{r}
## October 2013 course
causal2_ICPM = tab2_ICPM["causal",]
totals2_ICPM = addmargins(tab2_ICPM)["Sum",][1:2]
prop.test(causal2_ICPM, totals2_ICPM)
```

### Session Info

```{r}
session_info()
```
